---
title: "makeCodebook"
author: "vk"
date: "6/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Code Book for the wearables project

This code book describes the data characteristics of the project. More details on the data collection details are given at:
http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones

Here are the data for the project:
https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip.

## Data details:

## Subject id: 1- 30

## Activity labels: numeric: 1- 6. Corresponding activities are shown below.

 1 WALKING
 
 2 WALKING_UPSTAIRS
 
 3 WALKING_DOWNSTAIRS
 
 4 SITTING
 
 5 STANDING
 
 6 LAYING

## Features


Feature Selection 
=================

The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ. These time domain signals (prefix 't' to denote time) were captured at a constant rate of 50 Hz. Then they were filtered using a median filter and a 3rd order low pass Butterworth filter with a corner frequency of 20 Hz to remove noise. Similarly, the acceleration signal was then separated into body and gravity acceleration signals (tBodyAcc-XYZ and tGravityAcc-XYZ) using another low pass Butterworth filter with a corner frequency of 0.3 Hz. 

Subsequently, the body linear acceleration and angular velocity were derived in time to obtain Jerk signals (tBodyAccJerk-XYZ and tBodyGyroJerk-XYZ). Also the magnitude of these three-dimensional signals were calculated using the Euclidean norm (tBodyAccMag, tGravityAccMag, tBodyAccJerkMag, tBodyGyroMag, tBodyGyroJerkMag). 

Finally a Fast Fourier Transform (FFT) was applied to some of these signals producing fBodyAcc-XYZ, fBodyAccJerk-XYZ, fBodyGyro-XYZ, fBodyAccJerkMag, fBodyGyroMag, fBodyGyroJerkMag. (Note the 'f' to indicate frequency domain signals). 

These signals were used to estimate variables of the feature vector for each pattern:  
'-XYZ' is used to denote 3-axial signals in the X, Y and Z directions.


 tBodyAcc-XYZ
 
  tGravityAcc-XYZ

 tBodyAccJerk-XYZ

 tBodyGyro-XYZ

 tBodyGyroJerk-XYZ'

 tBodyAccMag

 tGravityAccMag

 tBodyAccJerkMag

 tBodyGyroMag

 tBodyGyroJerkMag

 fBodyAcc-XYZ

 fBodyAccJerk-XYZ

 fBodyGyro-XYZ

 fBodyAccMag

 fBodyAccJerkMag

 fBodyGyroMag

 fBodyGyroJerkMag

# The set of variables that were estimated from these signals are: 

 mean(): Mean value
 
 std(): Standard deviation
 
 mad(): Median absolute deviation 
 
 max(): Largest value in array
 
 min(): Smallest value in array
 
 sma(): Signal magnitude area
 
 energy(): Energy measure. Sum of the squares divided by the number of values.
 
 iqr(): Interquartile range 
 
 entropy(): Signal entropy
 
 arCoeff(): Autorregresion coefficients with Burg order equal to 4
 
 correlation(): correlation coefficient between two signals
 
 maxInds(): index of the frequency component with largest magnitude
 
 meanFreq(): Weighted average of the frequency components to obtain a mean frequency
 
 skewness(): skewness of the frequency domain signal
 
 kurtosis(): kurtosis of the frequency domain signal 
 
 bandsEnergy(): Energy of a frequency interval within the 64 bins of the FFT of each window.
 
 angle(): Angle between to vectors.

 Additional vectors obtained by averaging the signals in a signal window sample. These are used on the  angle() variable:

gravityMean

tBodyAccMean

tBodyAccJerkMean

tBodyGyroMean

tBodyGyroJerkMean

# Complete list of variables of each feature vector is available in 'features.txt'. For example:

1 tBodyAcc-mean()-X

2 tBodyAcc-mean()-Y

3 tBodyAcc-mean()-Z

4 tBodyAcc-std()-X

5 tBodyAcc-std()-Y

```{r}
make.codeBook <- function(dataSet) {
        df <- data.frame(
                row.names = NULL,
                column.names = names(dataSet),
                class = sapply(dataSet, class),
                range = sapply(dataSet, function(x)
                        if (class(x) == "factor")
                                paste(levels(x), collapse = " / ")
                        else if (class(x) == "numeric" ||
                                 class(x) == "integer")
                                paste(min(x), max(x), sep = "\t ")
                        else
                                class(x))
                
        )
        df
        
}
```
```{r}
uciDir <- "/Users/kumarnv/Rprojects/gettingData/projectGettingData/UCI HAR Dataset/"
trainDir <- "/Users/kumarnv/Rprojects/gettingData/projectGettingData/UCI HAR Dataset/train/"
setwd(uciDir)

featuresFile <- "features.txt" 
featuresData <- read.csv(featuresFile, sep="",header= FALSE)[2] # features names are in 2nd column, (561,1)
#activityLabelsFile <- list.files(uciDir)[1]
activityLabelsFile <- "activity_labels.txt" 
activityLabelsData <- read.csv(activityLabelsFile,sep="",header=FALSE) # dim 6,2

setwd(trainDir)
xTrainDataFile <- "X_train.txt"
yTrainDataFile <- "y_train.txt"
subjectTrainDataFile <- "subject_train.txt"

subjectTrainData <- read.csv(subjectTrainDataFile, sep="", header=FALSE) # 7352,1
names(subjectTrainData) <- "SubjectID"
xTrainData <- read.csv(xTrainDataFile, sep = "", header = FALSE) # 7352, 561
names(xTrainData) <- featuresData[,1]
yTrainData <- read.csv(yTrainDataFile, sep = "", header = FALSE) # labels 7352,1
names(yTrainData) <- "ActivityLabel"
# Merge subjectInfo, ActivityInfo & Sensor Data for training
mergedTrainData <- cbind(subjectTrainData,yTrainData,xTrainData) # 7352, 563
df <- make.codeBook(mergedTrainData)
setwd(uciDir)
write.table(df, "codeBook.md", sep = " \t ")
par <- read.table("codeBook.md")
par
```
